Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake

### 2014-10-19 paper.tex
Cover letter. Cleaned up bibliography.

@@

To further quantify the sensitivity of slip-deficit rate estimates to variability in data selection we perform a complete case resampling bootstrap shortening and slip-deficit rates are calculated for all possible subsets of the observed GPS velocities. There only $2^{20} = 1,048,576$ possible subsets of the observations (Figure \ref{fig:distribution}). We sample jointly from this distribution and the distribution of surface-area weighted fault depths. This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rates between 4 and 6 mm/yr. In the steepest, near-surface portion of the rangefront has minimum slip-deficit rates of ${\sim}$6 mm/yr, with an ${\sim}$25\% chance of slip-deficit rates above 10 mm/yr (Figure \ref{fig:distribution}). This geodetically constrained slip deficit rate can provide a mechanism for the rapid loading of the faults which exhibited 5-7 m of near-surface coseismic slip during the Wenchuan earthquake \citep{Xu2009, Shen2009a}.

@@

\bibliographystyle{plainnat}
\bibliography{biblio,library}
\end{document}

>>>>

\section{Cover Letter}
Dear Geology editors,

Please find attached ``Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake'' by T. Ben Thompson and Brendan J. Meade. We are submitting this article for consideration as an article in Geology. The manuscript contains 3249 words and 4 figures. (TBT Note: According to the Geology website, we are supposed to indicate whether we are willing to pay for color figures.)

The 2008 Wenchuan earthquake emphasized that the dicussion over tectonics in eastern Tibet has crucial implications for seismic hazard. Geodetic slip-deficit rates are a primary way of translating tectonic models into concrete hazard estimates. But, estimates are dependent on assumptions of fault geometry and consideration of earthquake cycle effects. Recent observations of co- and post-seismic slip on a deep detachment suggest that Longmen Shan regional fault geometry may be more complex than previously thought. Using a boundary element model, we analyze the implications of a interseismically locked detachment beneath eastern Tibet and show that geometry can produce a velocity gradient that is steepest 100-200 km northwest of the Longmen Shan rangefront, exactly as observed, with almost no velocity gradient at the rangefront. Inferring from GPS velocities measured before the earthquake, the resulting slip-deficit estimate is as high as 10 mm/yr at the steepest parts of the rangefront Beichuan fault. These results show a simple analysis of the GPS velocities across the Longmen Shan will produce dramatically different results than an analysis including more detailed fault geometry.

These results are of wider significance in that they demonstrate the impact of fault geometry on the analysis of fault systems, geodetic observation of a detachment participating in the earthquake cycle, and the effects of surface topography on geodesy. We are optimistic that these results will contribute to the understanding of the eastern Tibetan Plateau.  The active tectonics and seismic hazard of eastern Tibet and the Longmen Shan are addressed by other publications in Geology (Clark and Royden 2000, Meade 2007, OTHERS?), so the journal is a natural arena to continue the discussion.

Thank you for considering this paper for publication. We look forward to hearing from you.

T. Ben Thompson, Graduate Student

Department of Earth \& Planetary Sciences, Harvard University

Cambridge, MA 02138

(617)-NNN-NNNN, tthompson@fas.harvard.edu

@@

To further quantify the sensitivity of slip-deficit rate estimates to variability in data selection we perform a complete case resampling bootstrap shortening and slip-deficit rates are calculated for all possible subsets of the observed GPS velocities. There only $2^{20} = 1,048,576$ possible subsets of the observations (Figure \ref{fig:distribution}). We sample jointly from this distribution and the distribution of surface-area weighted fault depths. This analysis demonstrates that the majority of the slip-deficit-depth probability landscape lies at slip-deficit rates between 4 and 6 mm/yr. In the steepest, near-surface portion of the rangefront has minimum slip-deficit rates of ${\sim}$6 mm/yr, with an ${\sim}$25\% chance of slip-deficit rates above 10 mm/yr (Figure \ref{fig:distribution}). This geodetically constrained slip deficit rate can provide a mechanism for the rapid loading of the faults which exhibited 5-7 m of near-surface coseismic slip during the Wenchuan earthquake \citep{Xu2009, Shen2009b}.

@@

\bibliographystyle{Jcks}
\bibliography{biblio,library}
\end{document}
